### 2022-06-28 draft.tex
Deploying to gh-pages from @ PoisotLab/ms_metaweb_perspectives@1d38a4d0137defc303e636439c8f51544298a89c 🚀

\begin{enumerate}
    \item Metawebs, i.e.~networks of potential interactions within a
species pool, are a powerful abstraction to understand how large-scales
species interaction networks are structured.%
    \item Because metawebs are typically expressed at large spatial and
taxonomic scales, assembling them is a tedious and costly process;
predictive methods can help circumvent the limitations in data
deficiencies, by providing `draft' metawebs.%
    \item One way to improve the predictive ability is to maximize the
information used for prediction, by using graph embeddings rather than
the list of species interactions. Graph embedding is an emerging field
in machine learning that holds great potential for ecological problems.%
    \item In this perspective, we outline how the challenges associated
with inferring metawebs line-up with the advantages of graph embeddings;
furthermore, because metawebs are inherently spatial objects, we discuss
how the choice of the species pool has consequences on the reconstructed
network, but also embeds hypotheses about which human-made boundaries
are ecologically meaningful.%
\end{enumerate}

@@

Being able to infer \emph{potential} interactions could be the catalyst
for significant breakthroughs in our ability to start thinking about
species interaction networks over large spatial scales (Hortal et al.,
2015). Understanding species interactions holds enormous potential to
not only understand and more rapidly learn about species interactions
and metawebs, but also how changes in management of a single species may
impact non-target species. In a recent overview of the field of
ecological network prediction, Strydom, Catchen, et al. (2021)
identified two challenges of interest to the prediction of interactions
at large scales. First, there is a relative scarcity of relevant data in
most places globally -- paradoxically, this restricts our ability to
infer interactions for locations where inference is perhaps the least
required (and leaves us unable to make inference in regions without
interaction data); second, accurate predictions often demand accurate
predictors, and the lack of methods that can leverage small amount of
\emph{accurate} data is a serious impediment to our global predictive
ability. In most places, our most reliable biodiversity knowledge is
that of a species pool (\emph{i.e.} a set of potentially interacting
species in a given area): through the analysis of databases like GBIF or
IUCN, it is possible to construct a list of species in a region of
interest; but inferring the potential interactions between these species
is difficult.

Following the definition of Dunne (2006), a metaweb is the ecological
network analogue to the species pool; specifically, it inventories all
\emph{potential} interactions between species for a spatially delimited
area (and so captures the \(\gamma\) diversity of interactions). The
metaweb is not a prediction of the network at a specific point within
the spatial area it covers: it will have a different structure, notably
by having a larger connectance (see \emph{e.g.} Wood et al., 2015) and
complexity (see \emph{e.g.} Galiana et al., 2022), from any of these
local networks. These local networks (which capture the \(\alpha\)
diversity of interactions) are a subset of the metaweb's species and
their interactions, and have been called ``metaweb realizations''
(Poisot et al., 2015). Differences between local networks and their
metawebs are due to chance, species abundance and co-occurrence, local
environmental conditions, and local distribution of functional traits,
among others. Yet, recent results by Saravia et al. (2021) strongly
suggest that the local realizations only respond weakly to local
conditions: instead, they reflect constraints inherited by the structure
of their metaweb. This establishes the metaweb structure as the core
goal of predictive network ecology, as it is a required information to
accurately produce downscaled, local predictions.

Because the metaweb represents the joint effect of functional,
phylogenetic, and macroecological processes (Morales-Castilla et al.,
2015), it holds valuable ecological information. Specifically, it is the
``upper bounds'' on what the composition of the local networks, given
the local species pool, can be (see \emph{e.g.} McLeod et al., 2021);
this information can help evaluate the ability of ecological assemblages
to withstand the effects of, for example, climate change (Fricke et al.,
2022). These local networks may be reconstructed given an appropriate
knowledge of local species composition and provide information on the
structure of food webs at finer spatial scales. This has been done for
example for tree-galler-parasitoid systems (Gravel et al., 2018), fish
trophic interactions (Albouy et al., 2019), tetrapod trophic
interactions (Braga et al., 2019; O'Connor et al., 2020), and crop-pest
networks (Grünig et al., 2020). In this contribution, we highlight the
power in viewing (and constructing) metawebs as \emph{probabilistic}
objects in the context of rare interactions, discuss how a family of
machine learning tools (graph embeddings and transfer learning) can be
used to overcome data limitations to metaweb inference, and highlight
how the use of metawebs introduces important questions for the field of
network ecology.

\hypertarget{the-metaweb-is-an-inherently-probabilistic-object}{%
\section{The metaweb is an inherently probabilistic
object}\label{the-metaweb-is-an-inherently-probabilistic-object}}

Treating interactions as probabilistic (as opposed to binary) events is
a more nuanced and realistic way to represent them. Dallas et al. (2017)
suggested that most links in ecological networks are cryptic,
\emph{i.e.} uncommon or hard to observe. This argument echoes Jordano
(2016): sampling ecological interactions is difficult because it
requires first the joint observation of two species, and then the
observation of their interaction. In addition, it is generally expected
that weak or rare links to be more prevalent in networks than common or
rare links (Csermely, 2004), compared to strong, persistent links; this
is notably the case in food chains, wherein many weaker links are key to
the stability of a system (Neutel et al., 2002). In the light of these
observations, we expect to see an over-representation of low-probability
interactions under a model that accurately predicts interaction
probabilities. Yet the original metaweb definition, and indeed most past
uses of metawebs, was based on the presence/absence of interactions.
Moving towards \emph{probabilistic} metawebs, by representing
interactions as Bernoulli events (see \emph{e.g.} Poisot et al., 2016),
offers the opportunity to weigh these rare interactions appropriately.
The inherent plasticity of interactions is important to capture: there
have been documented instances of food webs undergoing rapid
collapse/recovery cycles over short periods of time (\emph{e.g.}
Pedersen et al., 2017). These considerations emphasize why metaweb
predictions should focus on quantitative (preferentially probabilistic)
predictions; this should constrain the suite of appropriate models.

Yet it is important to recall that a metaweb is intended as a catalogue
of all potential interactions, which is then filtered (Morales-Castilla
et al., 2015). In a sense, that most ecological interactions are elusive
can call for a slightly different approach to sampling: once the common
interactions are documented, the effort required in documenting each
rare interaction will increase exponentially. Recent proposals suggest
that machine learning algorithms, in these situations, can act as data
generators (Hoffmann et al., 2019): high quality observational data can
generate the core rules underpinning the network structure, and be
supplemented with synthetic data coming from predictive models,
increasing the volume of information available for inference. Indeed,
Strydom, Catchen, et al. (2021) suggested that knowing the metaweb may
render the prediction of local networks easier, because it fixes an
``upper bound'' on which interactions can exist. In this context a
probabilistic metaweb represents an aggregation of informative priors on
the interactions, elusive information with the potential to boost our
predictive ability (Bartomeus et al., 2016).

\begin{figure}
\hypertarget{fig:embedding}{%
\centering
\includegraphics{figures/conceptual_2.png}
\caption{Overview of the embedding process. A network (\emph{A}),
represented here as its adjacency matrix, is converted into a
lower-dimensional object (\emph{B}) where nodes, subgraphs, or edges
have specific values (see tbl.~\ref{tbl:methods} for an overview of
methods and their use for species interactions). For the purposes of
prediction, this low-dimensional object encodes feature vectors for
\emph{e.g.} the nodes. Embedding also allows to visualize the structure
in the data differently (see fig.~\ref{fig:illustration}), much like
with a principal component analysis. From a low-dimensional feature
vector, it is possible to develop predictive approaches. Nodes in an
ecological network are usually species (\textbf{C}), for which we can
leverage phylogenetic relatedness (\emph{e.g.} Strydom et al., 2022) or
functional traits to fill the values of additional species we would like
to project in this space (here for nodes I, J, K, and L) from the
embedding of known species (here, nodes A, B, C, and D). Because
embeddings can be projected back to a graph, this allows us to
reconstruct a network with these new species (\textbf{D}). This entire
cycle constitutes an instance of transfer learning, where the transfered
information is the representation of graph \textbf{A} through its
embedding.}\label{fig:embedding}
}
\end{figure}

@@

Graph embedding (fig.~\ref{fig:embedding}) is a varied family of machine
learning techniques aiming to transform nodes and edges into a vector
space (Arsov \& Mirceva, 2019), usually of a lower dimension, whilst
maximally retaining key properties of the graph (Yan et al., 2005).
Ecological networks are an interesting candidate for the widespread
application of embeddings, as they tend to possess a shared structural
backbone (Bramon Mora et al., 2018), which hints at structural
invariants that can be revealed at lower dimensions. Indeed, food webs
are inherently low-dimensional objects, and can be adequately
represented with less than ten dimensions (Braga et al., 2019; Eklöf et
al., 2013). Simulation results by Botella et al. (2022) suggest that
there is no best method to identify architectural similarities between
networks, and that multiple approaches need to be tested and compared to
the network descriptor of interest. This matches previous, more general
results on graph embedding, which suggest that the choice of embedding
algorithm matters for the results (Goyal \& Ferrara, 2018). In
tbl.~\ref{tbl:methods}, we present a selection of common graph embedding
methods, alongside examples of their use to predict species
interactions; most of these methods rely either on linear algebra, or on
pseudo-random walks on graphs.

One prominent family of approaches we do not discuss in the present
manuscript is Graph Neural Networks (GNN; Zhou et al., 2020); GNN are,
in a sense, a method to embed a graph into a dense subspace, but belong
to the family of deep learning methods, which has its own set of
practices (see \emph{e.g.} Goodfellow et al., 2016). An important issue
with methods based on deep learning is that because their parameter
space is immense, the sample size of the data fed into them must be
similarly large (typically thousands of instances). This is a
requirement for the model to converge correctly during training, but
this assumption is unlikely to be met given the size of datasets about
currently available metawebs (or single time/location species
interaction networks). This data volume requirement is mostly absent
from the techniques we list below. Furthermore, GNN still have some
challenges related to their shallow structure, and concerns related to
scalability (see Gupta et al., 2021 for a review), which are mostly
absent from the methods listed in tbl.~\ref{tbl:methods}. Assuming that
the uptake of next-generation biomonitoring techniques does indeed
deliver larger datasets on species interactions (Bohan et al., 2017),
there is a potential for GNN to become an applicable
embedding/predictive technique in the coming years.

@@

\begin{longtable}[]{@{}llrr@{}}

@@

approaches, by time of publication, alongside examples of their use in
the prediction of species interactions. These methods have not yet been
routinely used to predict species interactions; most examples that we
identified were either statistical associations, or analogues to joint
species distribution models. \(^a\): statistical interactions; \(^b\):
joint-SDM-like approach.}\tabularnewline

@@

\begin{minipage}[b]{0.11\columnwidth}\raggedright

@@

\end{minipage} & \begin{minipage}[b]{0.30\columnwidth}\raggedright
Embedding approach\strut
\end{minipage} & \begin{minipage}[b]{0.16\columnwidth}\raggedleft

@@

\end{minipage} & \begin{minipage}[b]{0.32\columnwidth}\raggedleft
Application in species interactions\strut

@@

\begin{minipage}[t]{0.11\columnwidth}\raggedright

@@

\end{minipage} & \begin{minipage}[t]{0.30\columnwidth}\raggedright
nodes through statistical divergence\strut
\end{minipage} & \begin{minipage}[t]{0.16\columnwidth}\raggedleft

@@

\end{minipage} & \begin{minipage}[t]{0.32\columnwidth}\raggedleft
Cieslak et al. (2020) \(^a\)\strut

@@

\begin{minipage}[t]{0.11\columnwidth}\raggedright

@@

\end{minipage} & \begin{minipage}[t]{0.30\columnwidth}\raggedright
graph through SVD\strut
\end{minipage} & \begin{minipage}[t]{0.16\columnwidth}\raggedleft

@@

\end{minipage} & \begin{minipage}[t]{0.32\columnwidth}\raggedleft

@@

\begin{minipage}[t]{0.11\columnwidth}\raggedright

@@

\end{minipage} & \begin{minipage}[t]{0.30\columnwidth}\raggedright
graph walk\strut
\end{minipage} & \begin{minipage}[t]{0.16\columnwidth}\raggedleft

@@

\begin{minipage}[t]{0.11\columnwidth}\raggedright

@@

\end{minipage} & \begin{minipage}[t]{0.30\columnwidth}\raggedright
graph through PCA/SVD analogue\strut
\end{minipage} & \begin{minipage}[t]{0.16\columnwidth}\raggedleft

@@

\begin{minipage}[t]{0.11\columnwidth}\raggedright
\end{minipage} & \begin{minipage}[t]{0.30\columnwidth}\raggedright
nodes through statistical divergence\strut
\end{minipage} & \begin{minipage}[t]{0.16\columnwidth}\raggedleft
\end{minipage} & \begin{minipage}[t]{0.32\columnwidth}\raggedleft

@@

\begin{minipage}[t]{0.11\columnwidth}\raggedright
\end{minipage} & \begin{minipage}[t]{0.30\columnwidth}\raggedright
nodes through auto-encoding\strut
\end{minipage} & \begin{minipage}[t]{0.16\columnwidth}\raggedleft
\end{minipage} & \begin{minipage}[t]{0.32\columnwidth}\raggedleft
\begin{minipage}[t]{0.11\columnwidth}\raggedright
\end{minipage} & \begin{minipage}[t]{0.30\columnwidth}\raggedright
nodes embedding\strut
\end{minipage} & \begin{minipage}[t]{0.16\columnwidth}\raggedleft
\end{minipage} & \begin{minipage}[t]{0.32\columnwidth}\raggedleft
\begin{minipage}[t]{0.11\columnwidth}\raggedright
\end{minipage} & \begin{minipage}[t]{0.30\columnwidth}\raggedright
sub-graph embedding\strut
\end{minipage} & \begin{minipage}[t]{0.16\columnwidth}\raggedleft
\end{minipage} & \begin{minipage}[t]{0.32\columnwidth}\raggedleft
\begin{minipage}[t]{0.11\columnwidth}\raggedright
\end{minipage} & \begin{minipage}[t]{0.30\columnwidth}\raggedright
joint nodes embedding\strut
\end{minipage} & \begin{minipage}[t]{0.16\columnwidth}\raggedleft
\end{minipage} & \begin{minipage}[t]{0.32\columnwidth}\raggedleft
\begin{minipage}[t]{0.11\columnwidth}\raggedright
\end{minipage} & \begin{minipage}[t]{0.30\columnwidth}\raggedright
nodes through a meta-strategy\strut
\end{minipage} & \begin{minipage}[t]{0.16\columnwidth}\raggedleft
\end{minipage} & \begin{minipage}[t]{0.32\columnwidth}\raggedleft
\begin{minipage}[t]{0.11\columnwidth}\raggedright
\end{minipage} & \begin{minipage}[t]{0.30\columnwidth}\raggedright
graph embedding\strut
\end{minipage} & \begin{minipage}[t]{0.16\columnwidth}\raggedleft
\end{minipage} & \begin{minipage}[t]{0.32\columnwidth}\raggedleft
\begin{minipage}[t]{0.11\columnwidth}\raggedright

@@

\end{minipage} & \begin{minipage}[t]{0.30\columnwidth}\raggedright

@@

\end{minipage} & \begin{minipage}[t]{0.16\columnwidth}\raggedleft

@@

The popularity of graph embedding techniques in machine learning is more
than the search for structural invariants: graphs are discrete objects,
and machine learning techniques tend to handle continuous data better.
Bringing a sparse graph into a continuous, dense vector space (Xu, 2020)
opens up a broader variety of predictive algorithms, notably of the sort
that are able to predict events as probabilities (Murphy, 2022).
Furthermore, the projection of the graph itself is a representation that
can be learned; Runghen et al. (2021), for example, used a neural
network to learn the embedding of a network in which not all
interactions were known, based on nodes metadata. This example has many
parallels in ecology (see fig.~\ref{fig:embedding} \textbf{C}), in which
node metadata can be given by phylogeny or functional traits. Rather
than directly predicting biological rules (see \emph{e.g.} Pichler et
al., 2020 for an overview), which may be confounded by the sparse nature
of graph data, learning embeddings works in the low-dimensional space
that maximizes information about the network structure. This approach is
further justified by the observation, for example, that the
macro-evolutionary history of a network is adequately represented by
some graph embeddings (RDPG; see Dalla Riva \& Stouffer, 2016). In a
recent publication, Strydom et al. (2022) have used an embedding (based
on RDPG) to project a metaweb of trophic interactions between European
mammals, and transfered this information to mammals of Canada by using
the phylogenetic distance between related clades to infer the values in
the latent sub-space into which the metaweb was projected. By performing
the RDPG step on re-constructed value, this approach yields a
probabilistic trophic metaweb for mammals of Canada based on knowledge
of European species, despite a limited (\(\approx\) 5\%) taxonomic
overlap.

>>>>

\begin{enumerate}
    \item Metawebs, (networks of potential interactions within a species
pool) are a powerful abstraction to understand how large-scale species
interaction networks are structured.%
    \item Because metawebs are typically expressed at large spatial and
taxonomic scales, assembling them is a tedious and costly process;
predictive methods can help circumvent the limitations in data
deficiencies, by providing `draft' metawebs.%
    \item One way to improve our ability to predict metawebs is to
maximize available information by using graph embeddings, instead of the
list of species interactions. Graph embedding is an emerging field in
machine learning that holds great potential for ecological problems.%
    \item In this perspective, we outline how the challenges associated
with inferring metawebs line-up with the advantages of graph embeddings;
as well as discuss how the choice of the species pool has consequences
on the reconstructed network, but also embeds hypotheses about which
human-made boundaries are ecologically meaningful.%
\end{enumerate}

@@

Being able to infer \emph{potential} interactions could serve as a
significant breakthrough in our ability to start thinking about species
interaction networks over large spatial scales (Hortal et al., 2015).
Understanding species interactions holds enormous potential to not only
understand and more rapidly learn about species interactions and
metawebs, but also how changes in management of a single species may
impact non-target species. In a recent overview of the field of
ecological network prediction, Strydom, Catchen, et al. (2021)
identified two challenges of interest to the prediction of interactions
at large scales. First, there is a relative scarcity of relevant data in
most places globally -- paradoxically, this restricts our ability to
infer interactions for locations where inference is perhaps the least
required (and leaves us unable to make inference in regions without
interaction data); second, accurate predictors are important for
accurate predictions, and the lack of methods that can leverage a small
amount of \emph{accurate} data is a serious impediment to our predictive
ability. In most places, our most reliable biodiversity knowledge is
that of a species pool (\emph{i.e.} a set of potentially interacting
species in a given area): through the analysis of databases like GBIF or
IUCN, it is possible to construct a list of species in a region of
interest; but inferring the potential interactions between these species
is difficult.

Following the definition of Dunne (2006), a metaweb is the ecological
network analogue to the species pool; specifically, it inventories all
\emph{potential} interactions between species for a spatially delimited
area (and so captures the \(\gamma\) diversity of interactions). The
metaweb is not a prediction of the network at a specific point within
the spatial area it covers: it will have a different structure, notably
by having a larger connectance (see \emph{e.g.} Wood et al., 2015) and
complexity (see \emph{e.g.} Galiana et al., 2022), from any of these
local networks. These local networks (which capture the \(\alpha\)
diversity of interactions) are a subset of the metaweb's species and
realized interactions, and have been called ``metaweb realizations''
(Poisot et al., 2015). Differences between local networks and their
metawebs are due to chance, species abundance and co-occurrence, local
environmental conditions, and local distribution of functional traits,
among others. Yet, recent results by Saravia et al. (2021) strongly
suggest that the local (metaweb) realizations only respond weakly to
local conditions: instead, they reflect constraints inherited by the
structure of their metaweb. This establishes the metaweb structure as
the core goal of predictive network ecology, as it is a required
information to accurately produce downscaled, local predictions.

Because the metaweb represents the joint effect of functional,
phylogenetic, and macroecological processes (Morales-Castilla et al.,
2015), it holds valuable ecological information. Specifically, it
represents the ``upper bounds'' on what the composition of the local
networks, given a local species pool, can be (see \emph{e.g.} McLeod et
al., 2021); this information can help evaluate the ability of ecological
assemblages to withstand the effects of, for example, climate change
(Fricke et al., 2022). These local networks may be reconstructed given
an appropriate knowledge of local species composition and provide
information on the structure of food webs at finer spatial scales. This
has been done for example for tree-galler-parasitoid systems (Gravel et
al., 2018), fish trophic interactions (Albouy et al., 2019), tetrapod
trophic interactions (Braga et al., 2019; O'Connor et al., 2020), and
crop-pest networks (Grünig et al., 2020). In this contribution, we
highlight the power in viewing (and constructing) metawebs as
\emph{probabilistic} objects in the context of rare interactions,
discuss how a family of machine learning tools (graph embeddings and
transfer learning) can be used to overcome data limitations to metaweb
inference, and highlight how the use of metawebs introduces important
questions for the field of network ecology.

\hypertarget{a-metaweb-is-an-inherently-probabilistic-object}{%
\section{A metaweb is an inherently probabilistic
object}\label{a-metaweb-is-an-inherently-probabilistic-object}}

Treating interactions as probabilistic (as opposed to binary) events is
a more nuanced and realistic way to represent them. Dallas et al. (2017)
suggested that most links in ecological networks are cryptic,
\emph{i.e.} uncommon or hard to observe. This argument echoes Jordano
(2016): sampling ecological interactions is difficult because it
requires first the joint observation of two species, and then the
observation of their interaction. In addition, it is generally expected
weak or rare links to be more prevalent in networks than common or
strong links (Csermely, 2004), compared to strong, persistent links;
this is notably the case in food chains, wherein many weaker links are
key to the stability of a system (Neutel et al., 2002). In the light of
these observations, we expect to see an over-representation of
low-probability (rare) interactions under a model that accurately
predicts interaction probabilities. Yet the original metaweb definition,
and indeed most past uses of metawebs, was based on the presence/absence
of interactions. Moving towards \emph{probabilistic} metawebs, by
representing interactions as Bernoulli events (see \emph{e.g.} Poisot et
al., 2016), offers the opportunity to weigh these rare interactions
appropriately. The inherent plasticity of interactions is important to
capture: there have been documented instances of food webs undergoing
rapid collapse/recovery cycles over short periods of time (\emph{e.g.}
Pedersen et al., 2017). These considerations emphasize why metaweb
predictions should focus on quantitative (preferentially probabilistic)
predictions, and this should constrain the suite of appropriate models
used to predict them.

It is important to recall that a metaweb is intended as a catalogue of
all potential interactions, which is then filtered for a given
application (Morales-Castilla et al., 2015). In a sense, that most
ecological interactions are elusive can call for a slightly different
approach to sampling: once the common interactions are documented, the
effort required in documenting each rare interaction will increase
exponentially. Recent proposals suggest that machine learning
algorithms, in these situations, can act as data generators (Hoffmann et
al., 2019): high quality observational data can generate the core rules
underpinning the network structure, and be supplemented with synthetic
data coming from predictive models, increasing the volume of information
available for inference. Indeed, Strydom, Catchen, et al. (2021)
suggested that knowing the metaweb may render the prediction of local
networks easier, because it fixes an ``upper bound'' on which
interactions can exist. In this context, a probabilistic metaweb
represents an aggregation of informative priors on the interactions,
elusive information with the potential to boost our predictive ability
(Bartomeus et al., 2016).

\begin{figure}
\hypertarget{fig:embedding}{%
\centering
\includegraphics{figures/conceptual_2.png}
\caption{Overview of the embedding process. A network (\textbf{A}),
represented here as its adjacency matrix, is converted into a
lower-dimensional object (\textbf{B}) where nodes, subgraphs, or edges
have specific values (see tbl.~\ref{tbl:methods} for an overview of
methods and their use for species interactions). For the purposes of
prediction, this low-dimensional object encodes feature vectors for
\emph{e.g.} the nodes. Embedding also allows to visualize the structure
in the data differently (see fig.~\ref{fig:illustration}), much like
with a principal component analysis. From a low-dimensional feature
vector, it is possible to develop predictive approaches. Nodes in an
ecological network are usually species (\textbf{C}), for which we can
leverage phylogenetic relatedness (\emph{e.g.} Strydom et al., 2022) or
functional traits to fill the values of additional species we would like
to project in this space (here for nodes I, J, K, and L) from the
embedding of known species (here, nodes A, B, C, and D). Because
embeddings can be projected back to a graph, this allows us to
reconstruct a network with these new species (\textbf{D}). This entire
cycle constitutes an instance of transfer learning, where the transfered
information is the representation of graph \textbf{A} through its
embedding.}\label{fig:embedding}
}
\end{figure}

@@

Graph (or Network) embedding (fig.~\ref{fig:embedding}) is a family of
machine learning techniques, whose main task is to learn a mapping
function from a discrete graph to a continuous domain (Arsov \& Mirceva,
2019; Chami et al., 2022). Their main goal is to learn a low dimensional
vector representations for the nodes of the graph (embeddings), such
that key properties of the graph (e.g.~local or global structures) are
retained in the embedding space(Yan et al., 2005). Ecological networks
are an interesting candidate for the widespread application of
embeddings, as they tend to possess a shared structural backbone (Bramon
Mora et al., 2018), which hints at structural invariants that can be
revealed at lower dimensions. Indeed, food webs are inherently
low-dimensional objects, and can be adequately represented with less
than ten dimensions (Braga et al., 2019; Eklöf et al., 2013). Simulation
results by Botella et al. (2022) suggest that there is no best method to
identify architectural similarities between networks, and that multiple
approaches need to be tested and compared to the network descriptor of
interest. This matches previous, more general results on graph
embedding, which suggest that the choice of embedding algorithm matters
for the results (Goyal \& Ferrara, 2018). In tbl.~\ref{tbl:methods}, we
present a selection of common graph and node embedding methods,
alongside examples of their use to predict species interactions; most of
these methods rely either on linear algebra, or on pseudo-random walks
on graphs.

One prominent family of approaches we do not discuss in the present
manuscript is Graph Neural Networks (GNN; Zhou et al., 2020). GNN are,
in a sense, a method to embed a graph into a dense subspace, but belong
to the family of deep learning methods, which has its own set of
practices (see \emph{e.g.} Goodfellow et al., 2016). An important issue
with methods based on deep learning is that, because their parameter
space is immense, the sample size of the data fed into them must be
similarly large (typically thousands of instances). This is a
requirement for the model to converge correctly during training, but
this assumption is unlikely to be met given the size of datasets
currently available for metawebs (or single time/location species
interaction networks). This data volume requirement is mostly absent
from the techniques we list below. Furthermore, GNN still have some
challenges related to their shallow structure, and concerns related to
scalability (see Gupta et al., 2021 for a review), which are mostly
absent from the methods listed in tbl.~\ref{tbl:methods}. Assuming that
the uptake of next-generation biomonitoring techniques does indeed
deliver larger datasets on species interactions (Bohan et al., 2017),
there is a potential for GNN to become an applicable
embedding/predictive technique in the coming years.

@@

\begin{longtable}[]{@{}lllll@{}}

@@

approaches, by type of embedded objects, alongside examples of their use
in the prediction of species interactions. These methods have not yet
been routinely used to predict species interactions; most examples that
we identified were either statistical associations, or analogues to
joint species distribution models. \(^a\): statistical interactions;
\(^b\): joint-SDM-like approach. Note that the row for PCA also applies
to kernel/probabilistic PCA, which are variations on the more general
method of SVD. Note further that tSNE has been included because it is
frequently used to embed graphs, including of species
associations/interactions, despite not being strictly speaking, a graph
embedding technique (see \emph{e.g.} Chami et al., 2022)}\tabularnewline

@@

\begin{minipage}[b]{0.09\columnwidth}\raggedright

@@

\end{minipage} & \begin{minipage}[b]{0.11\columnwidth}\raggedright
Object\strut
\end{minipage} & \begin{minipage}[b]{0.23\columnwidth}\raggedright
Technique\strut
\end{minipage} & \begin{minipage}[b]{0.14\columnwidth}\raggedright

@@

\end{minipage} & \begin{minipage}[b]{0.29\columnwidth}\raggedright
Application\strut

@@

\begin{minipage}[t]{0.09\columnwidth}\raggedright

@@

\end{minipage} & \begin{minipage}[t]{0.11\columnwidth}\raggedright
nodes\strut
\end{minipage} & \begin{minipage}[t]{0.23\columnwidth}\raggedright
statistical divergence\strut
\end{minipage} & \begin{minipage}[t]{0.14\columnwidth}\raggedright

@@

\end{minipage} & \begin{minipage}[t]{0.29\columnwidth}\raggedright
Gibb et al. (2021); Cieslak et al. (2020) \(^a\)\strut

@@

\begin{minipage}[t]{0.09\columnwidth}\raggedright
\end{minipage} & \begin{minipage}[t]{0.11\columnwidth}\raggedright
nodes\strut
\end{minipage} & \begin{minipage}[t]{0.23\columnwidth}\raggedright
stochastic gradient descent\strut
\end{minipage} & \begin{minipage}[t]{0.14\columnwidth}\raggedright
\end{minipage} & \begin{minipage}[t]{0.29\columnwidth}\raggedright
\begin{minipage}[t]{0.09\columnwidth}\raggedright
\end{minipage} & \begin{minipage}[t]{0.11\columnwidth}\raggedright
nodes\strut
\end{minipage} & \begin{minipage}[t]{0.23\columnwidth}\raggedright
gradient descent\strut
\end{minipage} & \begin{minipage}[t]{0.14\columnwidth}\raggedright
\end{minipage} & \begin{minipage}[t]{0.29\columnwidth}\raggedright
\begin{minipage}[t]{0.09\columnwidth}\raggedright
\end{minipage} & \begin{minipage}[t]{0.11\columnwidth}\raggedright
nodes\strut
\end{minipage} & \begin{minipage}[t]{0.23\columnwidth}\raggedright
stochastic gradient descent\strut
\end{minipage} & \begin{minipage}[t]{0.14\columnwidth}\raggedright
\end{minipage} & \begin{minipage}[t]{0.29\columnwidth}\raggedright
\begin{minipage}[t]{0.09\columnwidth}\raggedright
\end{minipage} & \begin{minipage}[t]{0.11\columnwidth}\raggedright
nodes\strut
\end{minipage} & \begin{minipage}[t]{0.23\columnwidth}\raggedright
meta-strategy\strut
\end{minipage} & \begin{minipage}[t]{0.14\columnwidth}\raggedright
\end{minipage} & \begin{minipage}[t]{0.29\columnwidth}\raggedright
\begin{minipage}[t]{0.09\columnwidth}\raggedright
\end{minipage} & \begin{minipage}[t]{0.11\columnwidth}\raggedright
joint nodes\strut
\end{minipage} & \begin{minipage}[t]{0.23\columnwidth}\raggedright
deep neural network\strut
\end{minipage} & \begin{minipage}[t]{0.14\columnwidth}\raggedright
\end{minipage} & \begin{minipage}[t]{0.29\columnwidth}\raggedright
\begin{minipage}[t]{0.09\columnwidth}\raggedright
\end{minipage} & \begin{minipage}[t]{0.11\columnwidth}\raggedright
sub-graph\strut
\end{minipage} & \begin{minipage}[t]{0.23\columnwidth}\raggedright
skipgram network\strut
\end{minipage} & \begin{minipage}[t]{0.14\columnwidth}\raggedright
\end{minipage} & \begin{minipage}[t]{0.29\columnwidth}\raggedright
\begin{minipage}[t]{0.09\columnwidth}\raggedright

@@

\end{minipage} & \begin{minipage}[t]{0.11\columnwidth}\raggedright
graph\strut
\end{minipage} & \begin{minipage}[t]{0.23\columnwidth}\raggedright
SVD\strut
\end{minipage} & \begin{minipage}[t]{0.14\columnwidth}\raggedright

@@

\end{minipage} & \begin{minipage}[t]{0.29\columnwidth}\raggedright

@@

\begin{minipage}[t]{0.09\columnwidth}\raggedright
GLEE\strut
\end{minipage} & \begin{minipage}[t]{0.11\columnwidth}\raggedright
graph\strut
\end{minipage} & \begin{minipage}[t]{0.23\columnwidth}\raggedright
Laplacian eigenmap\strut
\end{minipage} & \begin{minipage}[t]{0.14\columnwidth}\raggedright
(\textbf{Torres2020GleGeo?})\strut
\end{minipage} & \begin{minipage}[t]{0.29\columnwidth}\raggedright
\begin{minipage}[t]{0.09\columnwidth}\raggedright

@@

\end{minipage} & \begin{minipage}[t]{0.11\columnwidth}\raggedright
graph\strut
\end{minipage} & \begin{minipage}[t]{0.23\columnwidth}\raggedright
stochastic gradient descent\strut
\end{minipage} & \begin{minipage}[t]{0.14\columnwidth}\raggedright

@@

\begin{minipage}[t]{0.09\columnwidth}\raggedright
\end{minipage} & \begin{minipage}[t]{0.11\columnwidth}\raggedright
graph\strut
\end{minipage} & \begin{minipage}[t]{0.23\columnwidth}\raggedright
stochastic differential equation\strut
\end{minipage} & \begin{minipage}[t]{0.14\columnwidth}\raggedright
\end{minipage} & \begin{minipage}[t]{0.29\columnwidth}\raggedright
\begin{minipage}[t]{0.09\columnwidth}\raggedright

@@

\end{minipage} & \begin{minipage}[t]{0.11\columnwidth}\raggedright
graph\strut
\end{minipage} & \begin{minipage}[t]{0.23\columnwidth}\raggedright
eigen decomposition\strut
\end{minipage} & \begin{minipage}[t]{0.14\columnwidth}\raggedright

@@

\begin{minipage}[t]{0.09\columnwidth}\raggedright
PCA\strut
\end{minipage} & \begin{minipage}[t]{0.11\columnwidth}\raggedright
graph\strut
\end{minipage} & \begin{minipage}[t]{0.23\columnwidth}\raggedright
eigen decomposition\strut
\end{minipage} & \begin{minipage}[t]{0.14\columnwidth}\raggedright
(\textbf{S2013GraEmb?})\strut
\end{minipage} & \begin{minipage}[t]{0.29\columnwidth}\raggedright
Strydom, Catchen, et al. (2021)\strut

@@

\begin{minipage}[t]{0.09\columnwidth}\raggedright

@@

\end{minipage} & \begin{minipage}[t]{0.11\columnwidth}\raggedright

@@

\end{minipage} & \begin{minipage}[t]{0.23\columnwidth}\raggedright
multiple strategies\strut
\end{minipage} & \begin{minipage}[t]{0.14\columnwidth}\raggedright

@@

The popularity of graph embedding techniques in machine learning is more
than the search for structural invariants: graphs are discrete objects,
and machine learning techniques tend to handle continuous data better.
Bringing a sparse graph into a continuous, dense vector space (Xu, 2020)
opens up a broader variety of predictive algorithms, notably of the sort
that are able to predict events as probabilities (Murphy, 2022).
Furthermore, the projection of the graph itself is a representation that
can be learned; Runghen et al. (2021), for example, used a neural
network to learn the embedding of a network in which not all
interactions were known, based on the nodes' metadata. This example has
many parallels in ecology (see fig.~\ref{fig:embedding} \textbf{C}), in
which node metadata can be given by phylogeny or functional traits.
Rather than directly predicting biological rules (see \emph{e.g.}
Pichler et al., 2020 for an overview), which may be confounded by the
sparse nature of graph data, learning embeddings works in the
low-dimensional space that maximizes information about the network
structure. This approach is further justified by the observation, for
example, that the macro-evolutionary history of a network is adequately
represented by some graph embeddings (Random dot product graphs (RDPG);
see Dalla Riva \& Stouffer, 2016). In a recent publication, Strydom et
al. (2022) have used an embedding (based on RDPG) to project a metaweb
of trophic interactions between European mammals, and transferred this
information to mammals of Canada, using the phylogenetic distance
between related clades to infer the values in the latent sub-space into
which the European metaweb was projected. By performing the RDPG step on
re-constructed values, this approach yields a probabilistic trophic
metaweb for mammals of Canada based on knowledge of European species,
despite a limited (\(\approx\) 5\%) taxonomic overlap.
